Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2021-09-29 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd
@tyeentaylor , @m-n-smith -- one more thing to check

And @NidhiVinod (and @tyeentaylor ?)- should we add something on monoterpenes?

Emission of the VOC isoprene gas from tree leaves is a temperature-sensitive process associated with photosynthetic thermal tolerance [@taylorCapacityEmitIsoprene2019; @sharkeyIsopreneEmissionPlants2008], and within species, emission rates tend to increase toward brighter and hotter microenvironments [@niinemetsEmissionFactorVolatile2010]. 
Across landscapes, emitting species increase in relative abundance toward hotter climates, exceeding half of trees in warm tropical forests [@taylorIsopreneEmissionStructures2018a].
In contrast, vertical structuring of leaf isoprene emissions within forests may indicate the importance of temporal variability in thermal conditions as distinct from the long-term average.
The observation that both the abundance of emitting species (Table ) and average emission rates were highest in the mid-canopy region of an Amazonian forest [@taylorNewFieldInstrument2021] -- as opposed to the the hotter and brighter upper canopy-- is consistent with the hypothesis that temperature sensitive isoprene emissions enable real-time acclimation to rapidly changing leaf thermal environments typical of the mid-canopy region [see section 1, @sharkeyIsopreneEmissionPlants2008]. 
Future work seeking to understand how temporal dynamics of leaf heating [see @leighInfluenceLeafSize2017] vary through the canopy and influence leaf function will be important for resolving the role of isoprenes in forest thermal sensitivity across the vertical gradient.

@@

**Scaling in situ data with remote sensing**

@@

## Acknowledgements
Thanks to Norbert Kunert, Camille Piponiot, and Andrew Davies for providing figure materials. 
This manuscript benefited from feedback from Eleinis Ávila-Lovera, ... and the ForestGEO Ecosystems & Climate lab at SCBI.
Funding was provided by the Smithsonian Institution, ....

>>>>

Emission of the VOC isoprene gas from tree leaves is a temperature-sensitive process associated with photosynthetic thermal tolerance [@taylorCapacityEmitIsoprene2019; @sharkeyIsopreneEmissionPlants2008], and within species, emission rates tend to increase toward brighter and hotter microenvironments [@niinemetsEmissionFactorVolatile2010]. 
Across landscapes, emitting species increase in relative abundance toward hotter climates, exceeding half of trees in warm tropical forests [@taylorIsopreneEmissionStructures2018a].
In contrast, vertical structuring of leaf isoprene emissions within forests may indicate the importance of temporal variability in thermal conditions as distinct from the long-term average.
The observation that both the abundance of emitting species (Table ) and average emission rates were highest in the mid-canopy region of an Amazonian forest [@taylorNewFieldInstrument2021] -- as opposed to the the hotter and brighter upper canopy-- is consistent with the hypothesis that temperature sensitive isoprene emissions enable real-time acclimation to rapidly changing leaf thermal environments typical of the mid-canopy region [see section 1, @sharkeyIsopreneEmissionPlants2008]. 
Future work seeking to understand how temporal dynamics of leaf heating [see @leighInfluenceLeafSize2017] vary through the canopy and influence leaf function will be important for resolving the role of isoprenes in forest thermal sensitivity across the vertical gradient.
In the meantime, our best understannding is that isoprenes are important for mid-canopy thermal regulation, but that other traits are more important in the most sun-exposed canopy leaves.

*add something on monoterpenes?*

@@

### Scaling in situ data with remote sensing

@@

## Acknowledgements
Thanks to Norbert Kunert, Camille Piponiot, Peter B. Boucher and Andrew B. Davies for providing figure materials. 
This manuscript benefited from feedback from Eleinis Ávila-Lovera, ... and the ForestGEO Ecosystems & Climate lab at SCBI.
Funding was provided by the Smithsonian Institution, ....
